Journal and article coverage

### 2024-11-12 analysis/results.Rmd
Allow scp ta matching when there is more than one affiliation ID available

>>>>

## Replication

Here, I want to check whether we find consistent results across the three databases relative to open access uptake in hybrid journals in general, and relative to open access through agreements in particular. A particular focus will be whether findings change substantially when counting first or corresponding authors

Working definition replication: In scientific research, a replication is commonly defined as a study that is conducted using the same or similar methods as the original investigation, in order to evaluate whether consistent results can be obtained [1]. https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/

```{r}
# Web of Science
wos_jn_by_year <- readr::read_csv(here::here("data", "wos_jn_by_year.csv"))

wos_jn_ta_by_year <- readr::read_csv(here::here("data", "wos_jn_ta_by_year.csv"))

wos_ta_df <- wos_jn_by_year |>
  # only hybrid journals with at least one oa between 2018 - 2023
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  left_join(wos_jn_ta_by_year, by = c("issn_l", "earliest_year", "pub_type" = "pubtype")) |>
  # Focus on core
  filter(pub_type == "core") |>
  mutate(src = "Web of Science") |>
  select(-pub_type, -p)

# Scopus
scp_jn_by_year <-  readr::read_csv(here::here("data", "scp_jn_by_year.csv"))

scp_jn_ta_by_year <- readr::read_csv(here::here("data", "scp_ta_jn_by_year.csv"))

scp_ta_df <-  scp_jn_by_year |>
  # only hybrid journals with at least one oa between 2018 - 2023
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  left_join(scp_jn_ta_by_year, by = c("issn_l", "earliest_year" = "first_pubyear", "pub_type" = "pubtype")) |>
  # Focus on core
  filter(pub_type == "core") |>
  mutate(src = "Scopus") |>
  select(-pub_type, -p)

# HOAD
hoad_jn_by_year_ <- active_jns_with_oa |>
  mutate(cr_year = as.numeric(as.character(cr_year))) |>
  distinct(issn_l, cr_year, n_articles = jn_all)
hoad_oa_jn_by_year <- hoaddata::cc_articles |>
  group_by(issn_l, cr_year) |>
  summarise(n_oa_articles = n_distinct(doi))

hoad_jn_by_year <- inner_join(hoad_jn_by_year_, hoad_oa_jn_by_year, by = c("issn_l", "cr_year"))

hoad_jn_ta_by_year <- readr::read_csv(here::here("data-raw", "ta_oa_inst.csv"))

hoad_ta_by_year <- hoad_jn_ta_by_year |>
  filter(issn_l %in% active_jns_with_oa$issn_l, ta_active == TRUE) |>
  group_by(issn_l, cr_year) |>
  summarise(ta_first_author_articles = n_distinct(doi))

hoad_ta_oa_by_year <- hoad_jn_ta_by_year |>
  filter(issn_l %in% active_jns_with_oa$issn_l, ta_active == TRUE, !is.na(cc)) |>
  group_by(issn_l, cr_year) |>
  summarise(ta_oa_first_author_articles = n_distinct(doi))

hoad_jn_ta_by_year_ <- left_join(hoad_ta_by_year, hoad_ta_oa_by_year)

hoad_ta_df <- hoad_jn_by_year |>
  left_join(hoad_jn_ta_by_year_) |>
  # https://stackoverflow.com/a/73621978
  {\(.) {replace(.,is.na(.),0)}}() |>
  mutate(src = "HOAD") |>
  distinct(issn_l, earliest_year = cr_year, n_oa_articles, n_articles, ta_articles = ta_first_author_articles, ta_first_author_articles, ta_oa_articles = ta_oa_first_author_articles, ta_oa_first_author_articles, src)

uptake_df <- bind_rows(wos_ta_df, scp_ta_df, hoad_ta_df) |>
  mutate_if(is.numeric, ~replace(., is.na(.), 0))
```

```{r plot_uptake}
# Data prep
uptake_by_year <- uptake_df |>
  group_by(earliest_year, src) |>
  summarize(
    n_articles = sum(n_articles),
    n_oa_articles = sum(n_oa_articles),
    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
  )

uptake_by_year_df <- uptake_by_year |>
  # Remove HOAD because we only have first author data available
  filter(src != "HOAD") |>
  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))

hoad_panel <- uptake_by_year |>
  filter(src == "HOAD") |>
  ggplot(aes(x = earliest_year)) +
  geom_bar(
    aes(y = n_oa_articles / n_articles, fill = "All"),
    stat = "identity",
    color = "#000000"
  ) +
  geom_bar(
    aes(y = ta_oa_first_author_articles / n_articles, fill = "TA OA"),
    stat = "identity",
    color = "#000000"
  ) +
  geom_text(
    aes(
      label = paste0(
        round(ta_oa_first_author_articles / n_oa_articles * 100, 1),
        "%\n(",
        format(ta_oa_first_author_articles, big.mark = ","),
        ")"
      ),
      y = ta_oa_first_author_articles / n_articles),
    position = position_dodge(width = 1),
    vjust = -0.2,
    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
    size = 0.8 * my_base_size / .pt
  ) +
  scale_fill_manual(
    values = c("#b3b3b3a0", "#56B4E9"),
    name = "",
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    limits = c(0, .20),
    breaks = c(0, .05, .1, .15, .20)
  ) +
  theme_minimal(base_size = my_base_size) +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

wos_scp_uptake_panel <- uptake_by_year_df |>
  distinct(earliest_year, n_oa_articles, n_articles, src, name, value) |>
  # Modify factor levels to control order
  mutate(
    src = factor(src, levels = c("Scopus", "Web of Science")),
    name = factor(name, 
                 levels = c("ta_oa_first_author_articles", "ta_oa_corresponding_author_articles"))
  ) |>
  ggplot() +
  geom_bar(
    aes(
      x = earliest_year,
      y = n_oa_articles / n_articles,
      fill = "All"
    ),
    stat = "identity",
    color = "#000000"
  ) +
  geom_bar(
    aes(
      x = earliest_year,
      y = value / n_articles,
      fill = "TA OA"
    ),
    stat = "identity",
    color = "#000000"
  ) +
   geom_text(
    aes(
      label = paste0(
        round(value / n_oa_articles * 100, 1),
        "%\n(",
        format(value, big.mark = ","),
        ")"
      ),
      x = earliest_year,
      y = value / n_articles),
    position = position_dodge(width = 1),
    vjust = -0.2,
    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
    size = 0.8 * my_base_size / .pt
  ) +
  facet_grid(fct_rev(src) ~ name) +
  scale_fill_manual(
    values = c("#b3b3b3a0", "#56B4E9"),
    name = "",
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    limits = c(0, .20),
    breaks = c(0, .05, .1, .15, .20)
  ) +
  theme_minimal(base_size = my_base_size) +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank())
```
